Wireless smart monitoring energy system: a proof of concept

### 2021-03-18 paper/main.tex
corrections in system results and challenges

\section{System results}

\section{Consumption}
%Qui mettiamo workflow generale con immagine algoritmo

\section{Challenges}
During the development, we found some issues outside the scope of this POC. Since the focus of the project wasn't on security we decided for ease of debugging to do not encrypt the packets exchanged between plugs and server. The plugs configuration is hard-coded, for example the Wi-Fi SSID and password are stored as a char array in our C++ code. Since the EPS 8266 could switch to access point mode which means generate a new wireless network, a configuration procedure could be implemented. A device - such as a smartphone - could connect to the plug wireless network and send the SSID and password required by our home wireless settings. Those information should be stored in a persistent drive - as a micro SD card - and also, in order to not increase the POC complexity and keep the costs low, we did not implemented those procedures.
Another challenge to face is the security issue, the information send from a plug to the server are completely clear, so there is no encryption implemented. A man-in-the-middle attack could intercept those packages and change their contents. An attackers could also pose as smart plug, sending fake power readings to the server interfering with the normal operation.
\section{Conclusions}
In the current day and age, we have the opportunity to make smart the most mundane objects. Smart plugs can be used to monitor energy consumption, give control to the user, and help the user reach a maximum consumption goal. These plugs can reliably operate using the UDP protocol provided there are some reliability checks at the application level and can do it in a busy home WLAN where traffic can be both TCP and UDP.

>>>>

\section{System results}
We used Wireshark to monitor our network while running the system. Generally, TCP traffic worsens the performances of UDP in wireless applications. Therefore, during the test, we had both TCP and UDP third-party traffic, in particular, we had a 1080p Netflix streaming (UDP) and a download of a Ubuntu ISO (TCP). The main objective of the test was to check the efficacy of our version of reliable UDP, so the focus was on the reception of acknowledgments by the server. The packets are sent to the correct client on average within 3.3 ms.
 \section{Consumption}
%Qui mettiamo workflow generale con immagine algoritmo

\section{Challenges}
During the development, we found some issues outside the scope of this POC. Since the focus of the project wasn't on security we decided for ease of debugging to do not encrypt the packets exchanged between plugs and server. Another issue is that the configuration of the plug is hard-coded, for example, the Wi-Fi SSID and password are stored as a char array in our C++ code. Since the EPS 8266 could switch to access point mode which means generate a new wireless network, a configuration procedure could be implemented. A device - such as a smartphone - could connect to the plug wireless network and send the SSID and password required by our home wireless settings. That information should be stored in a persistent drive - like a micro SD card.
Another challenge to face is the security issue, the information sent from a plug to the server is completely clear, so there is no encryption implemented. A man-in-the-middle attack could intercept those packets and change their contents. An attacker could also pose as a smart plug, sending fake power readings to the server interfering with the normal operation.

\section{Conclusions}
In the current day and age, we have the opportunity to make smart the most mundane objects. Smart plugs can be used to monitor energy consumption, give control to the user, and help the user reach a maximum consumption goal. These plugs can reliably operate using the UDP protocol provided there are some reliability checks at the application level and can do it in a busy home WLAN where traffic can be both TCP and UDP.
